---
title: "Microsoft Word and PowerPoint generation from R"
author: David Gohel - david.gohel@ardata.fr
date: 15/04/2016
output:
  ioslides_presentation:
    css: styles.css
runtime: knitr
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
```

```{r message=FALSE, echo=FALSE}
library(shiny)
library(ggplot2)
```

# ReporteRs | an R package for reporting automation

## Motivations

> Reporting automation from R

MS formats are still ubiquitous in corporate environments

Formats and complex layouts are strong requirements

Document updates and graphical annotations are time consuming

Deploiement must be easy 

## Solution 

> An R API for creating 

* MS documents 
* that look like corporate documents 
* with complex tables
* and editable graphics where specific annotations can be made by non R coders

## Example 

### Graphical and tabular objets to use (1)

```{r}
library(ReporteRs)
library(magrittr)
library(ggplot2)

myplot <- ggplot(data = mtcars,
                 aes(x = wt, y = qsec,
                     color = drat)) + geom_point()
  
my_ft <- vanilla.table(mtcars)
```

## Example 

### Graphical and tabular objets to use (2)


```{r}
myplot
```

## Example 

### Graphical and tabular objets to use (3)


```{r echo=FALSE, results='asis'}
tmp <- paste0("<style>table {border-collapse:collapse;}</style>", as.html(my_ft))
tmp <- URLencode(tmp)
cat('<iframe src="data:text/html;charset=utf-8,', tmp ,
    '" style="border: none; seamless:seamless; width: 800px; height: 600px"></iframe>')
```


## Word example


```{r}
docx() %>% 
  addTitle("Tabular example") %>% 
  addFlexTable( my_ft ) %>% 
  addTitle("Graphic example") %>% 
  addPlot(fun = print, x = myplot, vector.graphic = TRUE) %>% 
  writeDoc("docs/first_.docx")
```

It produces this [document](docs/first_.docx).


## PowerPoint example


```{r}
pptx() %>% 
  addSlide("Title and Content") %>% 
  addTitle("Tabular example") %>% 
  addFlexTable( my_ft ) %>% 
  addSlide("Title and Content") %>% 
  addTitle("Graphic example") %>% 
  addPlot(fun = print, x = myplot) %>% 
  writeDoc("docs/first_.pptx")
```

It produces this [document](docs/first_.pptx).

## Workflow

```{r, echo=FALSE}
DiagrammeR::grViz("
digraph dot{

graph [layout = circo]

node [shape = square,
      style = filled, 
      fillcolor = '#D46A6A']

a [label = 'object creation']
b [label = 'add content']
c [label = 'write document']

edge [color = grey, arrowhead = vee]
a -> {b}
b -> {c}
}
")
```


## *add* functions

|              |              |
|--------------|--------------|
| addFlexTable | addPlot      |
| addParagraph | addTitle     |
| addTOC       | addBreakPage |
| addSlide     | addSection   |
| addImage     | addMardown   |
| ...          | ...          |







# FlexTable | build complex formatted tables 

## FlexTable features


* Layout
    * Free table header and footer
    * Merge cells (by rows or columns)
* Format : cells, borders, paragraphs, text
* Annotate : add text and images 

## FlexTable output

```{r echo=FALSE, results='asis'}
# correlation matrix ------------------------------------------------------
data( mtcars )
cormatrix = cor(mtcars)

# color calculation -------------------------------------------------------
library( RColorBrewer )
colors = c("#FFFFFF", brewer.pal(n = 4, name = "Reds") )
breaks = c( -Inf, seq(0.2, .8, by = .2), Inf )
mycut = cut( abs(cormatrix), breaks = breaks, include.lowest = TRUE, label = FALSE )
mycolors = colors[ mycut ]

# FlexTable creation -------------------------------------------------------
corr_table = FlexTable( round(cormatrix, 1), add.rownames = TRUE,
  body.par.props = parCenter(padding=3), header.par.props = parCenter(padding=3) )
corr_table = setFlexTableBackgroundColors( corr_table, j = seq_len(ncol(cormatrix)) + 1, colors = mycolors )


# cosmetics ---------------------------------------------------------------
corr_table = setFlexTableBackgroundColors( corr_table, i = 1, colors = "#EFEFEF", to = "header" )
corr_table = setFlexTableBackgroundColors( corr_table, j = 1, colors = "#EFEFEF" )
corr_table[,1] = textBold()
tmp <- paste0("<style>table {border-collapse:collapse;}</style>", as.html(corr_table))
tmp <- URLencode(tmp)

cat('<iframe src="data:text/html;charset=utf-8,', tmp ,
    '" style="border: none; seamless:seamless; width: 800px; height: 600px"></iframe>')
```

## FlexTable API

* Documentated here : http://davidgohel.github.io/ReporteRs/FlexTable.html
* Sugar functions are available in package `rtable`.

## FlexTable example | data

```{r echo=FALSE, message=FALSE, warning=FALSE}
library( boot )
library( dplyr )

data(melanoma)

melanoma$Status = c( 'Melanoma', 'Alive' , 'Non-melanoma' )[melanoma$status]
melanoma$Gender = ifelse( melanoma$sex > 0, 'Male', 'Female' )
melanoma$Ulceration = ifelse( melanoma$ulcer > 0, 'Present', 'Absent' )

# data summary ------------------------------------------------------------
data = melanoma %>% group_by( Status, Gender, Ulceration ) %>% 
  summarize( n = n(),
             mean = round( mean( thickness, na.rm = T ), 3 ), 
             sd = round( sd( thickness, na.rm = T ), 3 ) )
data
```

## FlexTable example | layout (1)

```{r}
my_flextable = FlexTable( data ) %>%
  spanFlexTableRows(j = 1, runs = data$Status) %>% 
  spanFlexTableRows(j = 2, runs = data$Gender) %>% 
  addHeaderRow( value = c("Group by", "Statistics"), 
                colspan = c(3, 3), first = T ) %>% 
  setFlexTableWidths(widths = c(1.5, 1.5, 1.5, .75, .75, .75))
```

## FlexTable example | layout (2)

```{r echo=FALSE, results='asis'}
tmp <- paste0("<style>table {border-collapse:collapse;}</style>", as.html(my_flextable))
tmp <- URLencode(tmp)
cat('<iframe src="data:text/html;charset=utf-8,', tmp ,
    '" style="border: none; seamless:seamless; width: 800px; height: 600px"></iframe>')
```

## FlexTable example | format (1)

```{r}
my_flextable[to = "header"] = textBold(font.size=12)
my_flextable[to = "header"] = parCenter(padding=2)

my_flextable[] = textNormal(font.size=11)

my_flextable[data$mean < 2, "mean" ] = 
  textItalic(color = "gray", font.size=11)
my_flextable[data$mean > 3, "mean" ] = 
  textBold(color = "red", font.size=11)

my_flextable[] = parCenter(padding=2)
```

## FlexTable example | format (2)

```{r echo=FALSE, results='asis'}
tmp <- paste0("<style>table {border-collapse:collapse;}</style>", as.html(my_flextable))
tmp <- URLencode(tmp)
cat('<iframe src="data:text/html;charset=utf-8,', tmp ,
    '" style="border: none; seamless:seamless; width: 800px; height: 600px"></iframe>')
```

## FlexTable example | annotate (1)

```{r}
my_flextable[data$sd > 2, "sd", 
             text.properties = textProperties(
               vertical.align = "superscript") ] = "*"
my_flextable <- addFooterRow(my_flextable, 
       value = "*", colspan = 6, 
       text.properties = textProperties(
         vertical.align = "superscript") )
my_flextable[1, 1, to = "footer"] = 
  "standard deviation is > 2"
```

## FlexTable example | annotate (2)

```{r echo=FALSE, results='asis'}
tmp <- paste0("<style>table {border-collapse:collapse;}</style>", as.html(my_flextable))
tmp <- URLencode(tmp)
cat('<iframe src="data:text/html;charset=utf-8,', tmp ,
    '" style="border: none; seamless:seamless; width: 800px; height: 600px"></iframe>')
```











# Editable Vector Graphics 

## Features

<div class="columns-2">

### Vector graphics

* high quality rendering
* scaled without loss of image resolution 

### Editable

* Graphics can be annotated 
* Modifications can be disabled

</div>

## Function `addPlot`

* Use option `vector.graphics=TRUE` 

```{r}
pptx() %>% addSlide("Title and Content") %>% 
  addPlot(fun = print, x = myplot, 
    vector.graphic = TRUE) %>% 
  writeDoc("docs/evg_.pptx")

docx() %>% addPlot(fun = print, 
    x = myplot, vector.graphic = TRUE) %>% 
  writeDoc("docs/evg_.docx")
```

### Outputs 

[evg_.pptx](docs/evg_.pptx) and [evg_.docx](docs/evg_.docx) .


# Ressources and informations

## Get packages

```{r eval=FALSE}
install.packages(c( "ReporteRs", "rtable") )
```

`rtable` is complementing `ReporteRs` for tabular reporting.

## Support 

Most common issues : 

* R 32 bits and java 64 bits 
* R 64 bits and java 32 bits 

### Getting help 

stackoverflow and tag `reporters`: http://stackoverflow.com/questions/tagged/reporters

google-groups: https://groups.google.com/forum/#!forum/reporters-package

### Issues - bugs

Github: https://github.com/davidgohel/ReporteRs/issues


## Tutorials

> Creating PowerPoint presentations 

http://www.sthda.com/english/wiki/create-and-format-powerpoint-documents-from-r-software

> Creating Word documents 

http://www.sthda.com/english/wiki/create-and-format-word-documents-using-r-software-and-reporters-package


# Live demo

## Demos

* Clinical reporting
    * for R users
    * shiny app for non R users
* Financial reporting

